---
title: "Assignment2"
author: "Tara"
date: '2020-03-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Import libraries and dataset
```{r}
library(rms)
library("lattice")
library("tidyverse")
library(kableExtra)
library(stringr)
load("~/Downloads/logitdata.RData")
assigndat

assigndat.dd <- datadist(assigndat)
options(datadist="assigndat.dd")

assigndat.noNA.dd <- datadist(assigndat.noNA)
options(datadist="assigndat.noNA.dd")

```


#Data Cleaning 
```{r}
#make event into a factor
assigndat$event <- as.factor(assigndat$event)
#remove missing variables, no imputations for this assignment
assigndat.noNA <- na.omit(assigndat)
```

#descriptive statistics for continuous variables
```{r}
#find which columns have NAs
colNA<- assigndat %>% 
  select(where(~ any(is.na(.)))) %>% 
  is.na() %>% 
  summary() 
 
#find proportion of NA 
proportionNA<- as.numeric(str_sub(colNA[3,], start=7, end=8))/nrow(assigndat)

#Rename vector
names(proportionNA) <- colnames(colNA)

```

```{r}
#Find median, mean, min, and mix for all continuous variables
sum.age<- summary(assigndat.noNA$age)[c(3:4, 1, 6)]
sum.bmi<- summary(assigndat.noNA$bmi)[c(3:4, 1, 6)]
sum.wt<- summary(assigndat.noNA$wt)[c(3:4, 1, 6)]
sum.ht<- summary(assigndat.noNA$ht)[c(3:4, 1, 6)]


dt1<- rbind(sum.age, sum.bmi,sum.wt, sum.ht)

stdev <- c(sd(assigndat.noNA$age),sd(assigndat.noNA$bmi),sd(assigndat.noNA$wt), sd(assigndat.noNA$ht))

permissing <- c(0, 16/2804*100, 10/2804*100, 10/2804*100, 10/2804*100)

dt2 <- cbind(dt1,stdev, permissing)

dt2 <- as.data.frame(dt2)
rownames(dt2) <- c("Age", "BMI", "Weight", "Height")
dt2<- rownames_to_column(dt2)

colnames(dt2) <- c("Variable", "Median", "Mean", "Min", "Max", "Standard Deviation", "percent missing data before removing NA (%)")

dt2 %>% 
  kable() %>%  
  kable_styling(full_width = T) 

```

to examine characteristic of study cohort look at mean and ad for continuous, frequency for categorical variables, report proportion of rising data if missing observations. Examined for categorical how distributed by outcome. If we identify categorical variable with one category with fewer cases, might be removed from the model - only going to descriptor report these. 

#descriptive statistics for continous variables
```{r}
summary(assigndat)
zero <- c(summary(assigndat.noNA$event)[1],summary(assigndat.noNA$side)[1], summary(assigndat.noNA$sympt)[1], summary(assigndat.noNA$sex)[1], summary(assigndat.noNA$htn)[1], summary(assigndat.noNA$afib)[1], summary(assigndat.noNA$arrhyth)[1],summary(assigndat.noNA$vhd)[1], summary(assigndat.noNA$orgfail)[1], summary(assigndat.noNA$diab)[1], summary(assigndat.noNA$smk)[1], summary(assigndat.noNA$tia)[1], summary(assigndat.noNA$stroke)[1], summary(assigndat.noNA$cvd)[1],  summary(assigndat.noNA$cexam)[1],summary(assigndat.noNA$cabg)[1])

one <- c(summary(assigndat.noNA$event)[2],summary(assigndat.noNA$side)[2], summary(assigndat.noNA$sympt)[2], summary(assigndat.noNA$sex)[2], summary(assigndat.noNA$htn)[2], summary(assigndat.noNA$afib)[2], summary(assigndat.noNA$arrhyth)[2],summary(assigndat.noNA$vhd)[2], summary(assigndat.noNA$orgfail)[2], summary(assigndat.noNA$diab)[2], summary(assigndat.noNA$smk)[2], summary(assigndat.noNA$tia)[2], summary(assigndat.noNA$stroke)[2], summary(assigndat.noNA$cvd)[2],  summary(assigndat.noNA$cexam)[2],summary(assigndat.noNA$cabg)[2])

demographics<- cbind(zero, one)
rownames(demographics) <- c("Event (0=no, 1=yes)", "side(0=left, 1=right)", "symptomatic (0=no, 1=yes)", "sex (0=male, 1=female)", "hypertension (0=no, 1=yes)", "History of Atrial Fibrillation (0=no, 1=yes)", "History of other cardiac arrhythmia (0=no, 1=yes)", "History of valvular heart disease (0=no, 1=yes)", "History of other organ failure (0=no, 1=yes)", "Diabetes (0=no, 1=yes)", "Current Smoker (0=no, 1=yes)", "TIA history (0=no, 1=yes)","History of Stroke (0=no, 1=yes)" , "History of MI, CHF or angina (0=no, 1=yes)", "Abnormal cardiac exam (0=no, 1=yes)","History of CABG surgery (0=no, 1=yes)")

missing <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
permissing <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1/2804*100, 0)
demographics <- cbind(demographics,permissing)


demographics <- as.data.frame(demographics)
demographics <- rownames_to_column(demographics)

colnames(demographics) <- c("Variable", "0", "1", "percent missing data before removing NA (%)")




demographics %>% 
  kable() %>% 
	kable_styling(full_width = T) 

7/(2787)*100

```

#Exploratory data analysis
distributions of continous variables
```{r}
hist(assigndat.noNA$wt)
plot(assigndat.noNA$event)
hist(assigndat.noNA$ht)
hist(assigndat.noNA$bmi)
hist(assigndat.noNA$age)

#use xtbs
```

#assumptions
1. 

#Model
```{r}
str(assigndat)

fit1 <- lrm(event~wt*diab + smk + age + sex + htn + ht, data = assigndat.noNA)
summary(fit1)
summary(fit1, diab="Yes")

summary(fit1, diab="No")

summary(fit1, sex="female")
```
lrt for whole model vs empty model


```{r}
anova(fit1)
```
in fit1: R^2 is generalization of R^2, 3 gs are different version of g=index on different scale, g is natural of model (predicted values logodds scale), gr on odds scale, gp on probability scale, brier. care more about how these are effected by the validation. 

These all change on model validation

“The odds of the outcome for males is 1.61 times that for females (95% CI of 1.25 to 2.08) controlled for blood pressure, age and cholesterol.”

The main effect of weigh in non-diabetics

The odds of the outcome for weight in non-diabetics is 0.707120 times for every 19.5kg increase in weight (95% CI of  0.525310 to 0.951840).


#plots
```{r}
plot(Predict(fit1,wt,diab, fun = plogis), ylab = "Probability")
plot(Predict(fit1,wt,diab))
plot(Predict(fit1, wt,diab, fun = exp), xlim = c(75,129), ylab = "Odds", main = "Figure 2. Predicted Odds Ratio by Weight (kg)")
```

#likelihood ratio test to test the interaction
```{r}
fit2 <- lrm(event~wt + diab + smk + age + sex + htn + ht, data = assigndat.noNA)

lrfit <- lrtest(fit2, fit1)
```
keep interaction in the model



#model validation
goodness of fit cindex, influential observations, overfitting
```{r}
#c-index
```
cindex is the proportion of pairs of subject where one has the outcome and the other does not, higher for subject wiht outcome probability
ROC curve
0.8 is a good discrimination
use Cxy rank for validation
Dxy/2+0.5
if Dxy=0 is no better than flipping coin, 1 is perfect
Brier score (B)- reliability, bigger B more unreliable
Hosmer-Lemeshow test - reliability X^2HL

```{r}
#c=index
#bootstrap validation
set.seed(1500)
#lrm(event~wt*diab + smk + age + sex + htn + ht ,data = assigndat.noNA,  x=TRUE, y=TRUE)

validate(fit1,B=300,bw=TRUE,type="individual",method=".632")

show.influence(which.influence(fit1, 0.2),data.frame(assigndat.noNA,dffits=resid(fit1,"dffits")),report=c("dffits"))

show.influence(which.influence(fit1, 0.2), data.frame(assigndat.noNA, dffits = resid(fit1), report = c("diffits")))

cindex <- 0.1886/2+0.5
```
consider Dxy, slope, R^2, brier for assignment
model calibration - calibration allows you to get predicted values NOT NEEDED FOR ASSIGNMENT
dfitts and dfbetas for logistic regression
threshold for dffits

```{r}
cindex.o <- (0.0380/2)+0.5
cindex.n <- (0.1885/2)+0.5
```


```{r}
fit1 <- update(fit1,x=TRUE,y=TRUE)

fit.dffits <- resid(fit1,type="dffits")
fit.dfbetas <- resid(fit1,type="dfbetas")
fit.hat <- resid(fit1,type="hat")

plot(fit.dffits)
#about predicted values yhats
#x axis is subject number, based on picture of it, where can i draw line to identify likely candidates to look at. All doing, is there anything that are influences. y axis is like standard deviation units. 0.4 to 0.5 range might be issues. 1.0 is a severe problem. greater than 0.2 might be issues.

#rows with less than -0.2
assigndat.noNA[fit.dffits < -0.2,]
#rows 48 and 84
#there are two influencers

#sensitivity analysis - refit without them

boxplot(fit.dfbetas)
show.influence(which.influence(fit1),assigndat.noNA)
#identify observations who's dfbetas exceed a particular threshold, default will be 0.2 dfbetas exceeding 0.2 in absolute value
#extract rows in show.influence
#count shows how many coefficients this observation had large dfbetta on
# * are the variables with large dfbetta
# 48, 73, 151, 174, 454

plot(fit.hat)
fithatdata <- assigndat.noNA[fit.hat > 0.04,]
fithatdata<- rownames_to_column(fithatdata)

fithatdata$rowname
#8, 14, 43, 48, 77, 84, 163, 174, 282, 311, 369, 381, 391, 415, 463
#73, 151, 454
```

partial residual plots
useless for binary variables
useful for continuous variables want straight line, want linear representation
- prime continous variables want straight line if nonlinearity there is an issue

```{r}
par(mfrow=c(1,2))
resid(fit1,"partial",pl='loess')
par(mfrow=c(2,3))

resid(eg2.sexage,"partial",pl='loess')
par(mfrow=c(1,3))
resid(eg3.sexage,"partial",pl='loess')
par(mfrow=c(1,1))


partial, pl 
```

